---
title: "Naturalness Analyses"
author: "Elizabeth Gaillard"
date: "2026-01-19"
output: html_document
---

Analysis of naturalness data as described in “Reliable and Valid Rating Data in Less Time with the Fast Image Rating Experiment” by Elizabeth P. Gaillard, Nakwon Rim, Kimberly L. Meidenbauer, Kyoung Whan Choe, and Marc G. Berman

# load packages
```{r}
library(lme4)
library(emmeans)
library(tidyverse)
```

# load helper functions
```{r}
source("fire_helper_utils.r")
```

# clean and load data
```{r}
# clean and load data
source("fire_data_cleaning.r")

# remove artifacts of cleaning process from environment
# keep relevant cleaned dataframes and participant counts per dataframe
# dataframes with "_final" in the name have had attention check images removed (see fire_data_cleaning.r)

keep <- c("parsed_likert_pref_final","parsed_likert_nat_final","parsed_slider_pref_final","parsed_slider_nat_final","parsed_pair_pref_final","parsed_pair_nat_final","parsed_fire_pref_final","parsed_fire_nat_final","parsed_likert_pref","parsed_likert_nat","parsed_slider_pref","parsed_slider_nat","parsed_pair_pref","parsed_pair_nat","parsed_fire_pref","parsed_fire_nat","likert_pref_n","likert_nat_n","slider_pref_n","slider_nat_n","pair_pref_n","pair_nat_n","fire_pref_n","fire_nat_n","preproc_prolific","parse_ratings","remove_attn_fail","split_clicks","calculate_mean_sd","remove_attn_trials","permutation_test","split_half_rel_all","n_subset_reliability","n_subset_correlation") 

rm(list = setdiff(ls(), keep))
```

# save csvs for use with simulation code
```{r}
# write.csv(parsed_slider_nat_final, "simulation_code/parsed_data/parsed_slider_nat_final.csv")
```

# select naturalness data
```{r}
# datasets without attention check images
parsed_likert_final <- parsed_likert_nat_final
parsed_slider_final <- parsed_slider_nat_final
parsed_pair_final <- parsed_pair_nat_final
parsed_fire_final <- parsed_fire_nat_final

# datasets with attention check images
parsed_likert <- parsed_likert_nat
parsed_slider <- parsed_slider_nat
parsed_pair <- parsed_pair_nat
parsed_fire <- parsed_fire_nat
```


# HYPOTHESIS 1: TIME
# A. PER TRIAL
## average time per trial with attention checks
```{r}
avg_time_per_trial <- calculate_mean_sd(parsed_likert, parsed_slider, parsed_pair, parsed_fire)

avg_time_per_trial <- avg_time_per_trial %>%
  mutate(n = c(nrow(parsed_likert),nrow(parsed_slider),nrow(parsed_pair),nrow(parsed_fire))) %>%
  mutate(SE = SD/sqrt(n)) %>%
  mutate(
    Mean = Mean/1000,
    SE = SE/1000
  )
```

## remove attention check trials
```{r}
likert_time_no_attn <- remove_attn_trials(parsed_likert, "image")
slider_time_no_attn <- remove_attn_trials(parsed_slider, "image")
pair_time_no_attn <- remove_attn_trials(parsed_pair, c("image11", "image12"))
fire_time_no_attn <- remove_attn_trials(parsed_fire, 
  c("image11", "image12", "image13", "image14",
    "image21", "image22", "image23", "image24",
    "image31", "image32", "image33", "image34"))
```

## average time per trial without attention checks
```{r}
avg_time_per_trial_no_attn <- calculate_mean_sd(likert_time_no_attn, slider_time_no_attn, pair_time_no_attn, fire_time_no_attn)

avg_time_per_trial_no_attn <- avg_time_per_trial_no_attn %>%
  mutate(n = c(nrow(likert_time_no_attn),nrow(slider_time_no_attn),nrow(pair_time_no_attn),nrow(fire_time_no_attn))) %>%
  mutate(SE = SD/sqrt(n)) %>%
  mutate(
    Mean = Mean/1000,
    SE = SE/1000
  )
```

## significance testing
### mixed-effects models
#### with attention checks
```{r}
parsed_likert$rating_method <- "Likert"
likert_time <- parsed_likert %>% select("participant_id","rt","rating_method")

parsed_slider$rating_method <- "Slider"
slider_time <- parsed_slider %>% select("participant_id","rt","rating_method")

parsed_pair$rating_method <- "Pair"
pair_time <- parsed_pair %>% select("participant_id","rt","rating_method")

parsed_fire$rating_method <- "FIRE"
fire_time <- parsed_fire %>% select("participant_id","rt","rating_method")

time_data <- rbind(likert_time, slider_time, pair_time, fire_time)

#mixed-effects model
model <- lmer(rt ~ rating_method + (1|participant_id), data = time_data)
emm <- emmeans(model, ~ rating_method)
pairs(emm)
```

#### without attention checks
```{r}
likert_time_no_attn$rating_method <- "Likert"
likert_time_no_attn <- likert_time_no_attn %>% select("participant_id", "rt", "rating_method")

slider_time_no_attn$rating_method <- "Slider"
slider_time_no_attn <- slider_time_no_attn %>% select("participant_id", "rt", "rating_method")

pair_time_no_attn$rating_method <- "Pair"
pair_time_no_attn <- pair_time_no_attn %>% select("participant_id", "rt", "rating_method")

fire_time_no_attn$rating_method <- "FIRE"
fire_time_no_attn <- fire_time_no_attn %>% select("participant_id", "rt", "rating_method")

time_data_no_attn <- rbind(likert_time_no_attn, slider_time_no_attn, pair_time_no_attn, fire_time_no_attn)

#mixed-effects model
model_no_attn <- lmer(rt ~ rating_method + (1|participant_id), data = time_data_no_attn)
emm <- emmeans(model_no_attn, ~ rating_method)
pairs(emm)
```

## plotting
### with attention checks
```{r}
boxplot(parsed_likert$rt/1000, parsed_slider$rt/1000, parsed_pair$rt/1000, parsed_fire$rt/1000,
  names = c("Likert", "Slider", "Pair", "FIRE"),
  xlab = "Method",
  ylab = "Avg. Time Per Trial (sec)",
  col = c("#15616d", "#001524", "#ff7d00", "#ffecd1"),
  border = "black", 
  notch = TRUE, 
  outline = FALSE,
  ylim = c(0, 40)
)
```

### without attention checks
```{r}
boxplot(likert_time_no_attn$rt/1000, slider_time_no_attn$rt/1000, pair_time_no_attn$rt/1000, fire_time_no_attn$rt/1000,
  names = c("Likert", "Slider", "Pair", "FIRE"),
  xlab = "Method",
  ylab = "Avg. Time Per Trial (sec)",
  col = c("#15616d", "#001524", "#ff7d00", "#ffecd1"),
  border = "black", 
  notch = TRUE, 
  outline = FALSE,
  ylim = c(0,40)
)
```

# B. PER PARTICIPANT
## average time per participant
### with attention checks
```{r}
#ssum trial times by participant with attention checks
likert_time <- parsed_likert %>% group_by(participant_id) %>% summarise(rt = sum(rt))
slider_time <- parsed_slider %>% group_by(participant_id) %>% summarise(rt = sum(rt))
pair_time <- parsed_pair %>% group_by(participant_id) %>% summarise(rt = sum(rt))
fire_time <- parsed_fire %>% group_by(participant_id) %>% summarise(rt = sum(rt))

#sdescriptive stats
avg_time_per_part <- calculate_mean_sd(likert_time, slider_time, pair_time, fire_time)

avg_time_per_part <- avg_time_per_part %>%
  mutate(n = c(320, 320, 320, 320)) %>%
  mutate(SE = SD / sqrt(n)) %>%
   mutate(
    Mean = Mean/60000,
    SD = SD/60000
  )
```

### without attention checks
```{r}
# sum trial times by participant without attention checks
likert_time_no_attn_part <- parsed_likert %>% 
  filter(!str_starts(image, "attn")) %>% 
  group_by(participant_id) %>% 
  summarise(rt = sum(rt))

slider_time_no_attn_part <- parsed_slider %>% 
  filter(!str_starts(image, "attn")) %>% 
  group_by(participant_id) %>% 
  summarise(rt = sum(rt))

pair_time_no_attn_part <- parsed_pair %>% 
  filter(!str_starts(image11, "attn") & !str_starts(image12, "attn")) %>% 
  group_by(participant_id) %>%
  summarise(rt = sum(rt))

fire_time_no_attn_part <- parsed_fire %>% 
  filter(!str_starts(image11, "attn") & !str_starts(image12, "attn") & !str_starts(image13, "attn")
         & !str_starts(image21, "attn") & !str_starts(image22, "attn") & !str_starts(image23, "attn")
         & !str_starts(image31, "attn") & !str_starts(image32, "attn") & !str_starts(image33, "attn")) %>% 
  group_by(participant_id) %>%
  summarise(rt = sum(rt))

#descriptive stats
avg_time_per_part_no_attn <- calculate_mean_sd(likert_time_no_attn, slider_time_no_attn, pair_time_no_attn, fire_time_no_attn)

avg_time_per_part_no_attn <- avg_time_per_part_no_attn %>%
  mutate(n = c(320, 320, 320, 320)) %>%
  mutate(SE = SD / sqrt(n)) %>%
   mutate(
    Mean = Mean/60000,
    SD = SD/60000
  )
```

## significance testing
### preprocessing
#### with attention checks
```{r}
likert_time$rating_method <- "Likert"
likert_time_part <- likert_time %>% select("participant_id", "rt", "rating_method")

slider_time$rating_method <- "Slider"
slider_time_part <- slider_time %>% select("participant_id", "rt", "rating_method")

pair_time$rating_method <- "Pair"
pair_time_part <- pair_time %>% select("participant_id", "rt", "rating_method")

fire_time$rating_method <- "FIRE"
fire_time_part <- fire_time %>% select("participant_id", "rt", "rating_method")

time_data_part <- rbind(likert_time_part, slider_time_part, pair_time_part, fire_time_part)

## convert to minutes
time_data_part <- time_data_part %>% mutate(rt = rt/60000)
```

#### without attention checks
```{r}
likert_time_no_attn$rating_method <- "Likert"
likert_time_part_no_attn <- likert_time_no_attn %>% select("participant_id", "rt", "rating_method")

slider_time_no_attn$rating_method <- "Slider"
slider_time_part_no_attn <- slider_time_no_attn %>% select("participant_id", "rt", "rating_method")

pair_time_no_attn$rating_method <- "Pair"
pair_time_part_no_attn <- pair_time_no_attn %>% select("participant_id", "rt", "rating_method")

fire_time_no_attn$rating_method <- "FIRE"
fire_time_part_no_attn <- fire_time_no_attn %>% select("participant_id", "rt", "rating_method")

time_data_part_no_attn <- rbind(likert_time_part_no_attn, slider_time_part_no_attn, pair_time_part_no_attn, fire_time_part_no_attn)

## convert to minutes
time_data_part_no_attn <- time_data_part_no_attn %>% mutate(rt = rt/60000)
```

### permutation testing 
#### with attention checks
```{r}
results <- list(
  likert_slider_part <- permutation_test(time_data_part, "Likert", "Slider"),
  likert_pair_part <- permutation_test(time_data_part, "Likert", "Pair"),
  likert_fire_part <- permutation_test(time_data_part, "Likert", "FIRE"),
  slider_pair_part <- permutation_test(time_data_part, "Slider", "Pair"),
  slider_fire_part <- permutation_test(time_data_part, "Slider", "FIRE"),
  pair_fire_part <- permutation_test(time_data_part, "Pair", "FIRE")
)


# get summary
summary <- do.call(rbind, lapply(results, function(res) {
  data.frame(
    method1 = res$method1,
    method2 = res$method2,
    observed_diff = res$observed_diff,
    p_value = res$p_value
  )
}))
```

#### without attention checks
```{r}
results_no_attn <- list(
  likert_slider_no_attn_part <- permutation_test(time_data_part_no_attn, "Likert", "Slider"),
  likert_pair_no_attn_part <- permutation_test(time_data_part_no_attn, "Likert", "Pair"),
  likert_fire_no_attn_part <- permutation_test(time_data_part_no_attn, "Likert", "FIRE"),
  slider_pair_no_attn_part <- permutation_test(time_data_part_no_attn, "Slider", "Pair"),
  slider_fire_no_attn_part <- permutation_test(time_data_part_no_attn, "Slider", "FIRE"),
  pair_fire_no_attn_part <- permutation_test(time_data_part_no_attn, "Pair", "FIRE")
)


#get summary
summary_no_attn <- do.call(rbind, lapply(results_no_attn, function(res) {
  data.frame(
    method1 = res$method1,
    method2 = res$method2,
    observed_diff = res$observed_diff,
    p_value = res$p_value
  )
}))
```

## plotting
### with attention checks
```{r}
boxplot(likert_time$rt/60000, slider_time$rt/60000, pair_time$rt/60000, fire_time$rt/60000,
  names = c("Likert", "Slider", "Pair", "FIRE"),
  xlab = "Method",
  ylab = "Avg. Time Per Participant (min)",
  col = c("#15616d", "#001524", "#ff7d00", "#ffecd1"),
  border = "black", 
  notch = TRUE, 
  outline = FALSE,
  ylim = c(0, 30)
)
```

### without attention checks
```{r}
boxplot(likert_time_no_attn_part$rt/60000, slider_time_no_attn_part$rt/60000, pair_time_no_attn_part$rt/60000, fire_time_no_attn_part$rt/60000,
  names = c("Likert", "Slider", "Pair", "FIRE"),
  main = "Preference",
  xlab = "Method",
  ylab = "Avg. Time Per Participant (min)",
  col = c("#15616d", "#001524", "#ff7d00", "#ffecd1"),
  border = "black", 
  notch = TRUE, 
  outline = FALSE,
  ylim = c(0, 30)
)
```


# HYPOTHESIS 2: RELIABILITY
# A. FULL PARTICIPANT SET
## split-half reliability
```{r}
split_likert <- split_half_rel_all(parsed_likert_final,320)
split_slider <- split_half_rel_all(parsed_slider_final,320)
split_pair <- split_half_rel_all(parsed_pair_final,320)
split_fire <- split_half_rel_all(parsed_fire_final,320)

reliability_likert <- split_likert$results
reliability_slider <- split_slider$results
reliability_pair <- split_pair$results
reliability_fire <- split_fire$results
```

## plot
```{r}
boxplot(reliability_likert, reliability_slider, reliability_pair, reliability_fire,
  names = c("Likert", "Slider", "Pair", "FIRE"),
  xlab = "Method",
  ylab = "Correlation Coefficient",
  col = c("#15616d", "#001524", "#ff7d00", "#ffecd1"),
  border = "black", 
  notch = TRUE, 
  outline = FALSE
)
```

# B. SUBSETS
## calculate reliability with samples of 50, 20, ...160 participants 
```{r}
n_values <- seq(100, 320, by = 10)

# likert
rel_likert <- data.frame(N = integer(), mean_rel_likert = double(), sd_rel_likert = double())
for (n in n_values) {
  rel <- n_subset_reliability(parsed_likert_final, n)
  rel_likert <- rbind(rel_likert, data.frame(N = n, 
                                       mean_rel_likert = rel$mean,
                                       sd_rel_likert = rel$sd))
}
# slider
rel_slider <- data.frame(N = integer(), mean_rel_slider = double(), sd_rel_slider = double())
for (n in n_values) {
  rel <- n_subset_reliability(parsed_slider_final, n)
  rel_slider <- rbind(rel_slider, data.frame(N = n, 
                                       mean_rel_slider = rel$mean,
                                       sd_rel_slider = rel$sd))
}
# pair
rel_pair <- data.frame(N = integer(), mean_rel_pair = double(), sd_rel_pair = double())
for (n in n_values) {
  rel <- n_subset_reliability(parsed_pair_final, n)
  rel_pair <- rbind(rel_pair, data.frame(N = n, 
                                       mean_rel_pair = rel$mean,
                                       sd_rel_pair = rel$sd))
}
# fire
rel_fire <- data.frame(N = integer(), mean_rel_fire = double(), sd_rel_fire = double())
for (n in n_values) {
  rel <- n_subset_reliability(parsed_fire_final, n)
  rel_fire <- rbind(rel_fire, data.frame(N = n, 
                                       mean_rel_fire = rel$mean,
                                       sd_rel_fire = rel$sd))
}
```

## reliability by number of trials and by time
```{r}
# add column for trial
rel_likert <- rel_likert %>% mutate(n_trials = N*106)
rel_slider <- rel_slider %>% mutate(n_trials = N*106)
rel_pair <- rel_pair %>% mutate(n_trials = N*68)
rel_fire <- rel_fire %>% mutate(n_trials = N*9)

# add column for time
rel_likert <- rel_likert %>% mutate(time = n_trials*(avg_time_per_trial_no_attn[1,2]))
rel_slider <- rel_slider %>% mutate(time = n_trials*(avg_time_per_trial_no_attn[2,2]))
rel_pair <- rel_pair %>% mutate(time = n_trials*(avg_time_per_trial_no_attn[3,2]))
rel_fire <- rel_fire %>% mutate(time = n_trials*(avg_time_per_trial[4,2]))
```

# HYPOTHESIS 3: VALIDITY
## correlate ratings collected via each method
```{r}
# average over participants
likert_avg <- parsed_likert_final %>% group_by(image) %>% summarise(mean_likert = mean(response))

slider_avg <- parsed_slider_final %>% group_by(image) %>% summarise(mean_slider = mean(response))

parsed_pair_final$response <- as.numeric(parsed_pair_final$response)
pair_avg <- parsed_pair_final %>% group_by(image) %>% summarise(mean_pair = mean(response))

parsed_fire_final$response <- as.numeric(parsed_fire_final$response)
fire_avg <- parsed_fire_final %>% group_by(image) %>% summarise(mean_fire = mean(response))

#join mean ratings to shared data frame
all_avg <- likert_avg %>%
  left_join(slider_avg, by = "image") %>%
  left_join(pair_avg, by = "image") %>%
  left_join(fire_avg, by = "image")
```

## correlation with increasing subsamples of participants
### correlations with mean image rating across all methods
```{r}
n_values <- seq(10, 320, by = 10)

#likert
corr_results_likert <- data.frame(N = integer(), mean_corr_likert = double(), sd_corr_likert = double())
for (n in n_values) {
  corr <- n_subset_correlation(parsed_likert_final, all_avg, n, "mean_likert")
  corr_results_likert <- rbind(corr_results_likert, data.frame(N = n, 
                                       mean_corr_likert = corr$mean,
                                       sd_corr_likert = corr$sd))
}

#slider
corr_results_slider <- data.frame(N = integer(), mean_corr_slider = double(), sd_corr_slider = double())
for (n in n_values) {
  corr <- n_subset_correlation(parsed_slider_final, all_avg, n, "mean_slider")
  corr_results_slider <- rbind(corr_results_slider, data.frame(N = n, 
                                       mean_corr_slider = corr$mean,
                                       sd_corr_slider = corr$sd))
}
#pair
corr_results_pair <- data.frame(N = integer(), mean_corr_pair = double(), sd_corr_pair = double())
for (n in n_values) {
  corr <- n_subset_correlation(parsed_pair_final, all_avg, n, "mean_pair")
  corr_results_pair <- rbind(corr_results_pair, data.frame(N = n, 
                                       mean_corr_pair = corr$mean,
                                       sd_corr_pair = corr$sd))
}
#fire
corr_results_fire <- data.frame(N = integer(), mean_corr_fire = double(), sd_corr_fire = double())
for (n in n_values) {
  corr <- n_subset_correlation(parsed_fire_final, all_avg, n, "mean_fire")
  corr_results_fire <- rbind(corr_results_fire, data.frame(N = n, 
                                       mean_corr_fire = corr$mean,
                                       sd_corr_fire = corr$sd))
}
```

```{r}

```

